---
title: "Differential expression"
output: 
  flexdashboard::flex_dashboard
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(ggp.rnaseq)
library(DESeq2)
library(DT)
library(crosstalk)
library(d3scatter)
library(plotly)
library(org.Mm.eg.db)
library(KEGGREST)
```

```{r include=FALSE}
comparison <- snakemake@wildcards[['comparison']]

params <- snakemake@config[['comparisons']][[comparison]]
bad.samples <- params[['bad-samples']]
new.design <- as.formula(params[['design']])
prefilter.string <- params[['prefilter']]
sample.subset <- params[['subset']]
de.attribute <- params[['attribute']]
de.group.1 <- params[['group1_plus']]
de.group.2 <- params[['group2_minus']]

alpha <- 0.1
```

```{r include=FALSE}
#Annoying difference in behavior between RMarkdown and R scripts when it comes to working directories...
res <- readRDS(paste0('../',snakemake@input[['rds']]))
#ggp.results(dds.filtered,c('genotype','KO','IL7Rcre'))
res %>% filter(!is.na(padj) & padj < 0.1) %>%
    mutate(baseMean=formatC(baseMean,format="e",digits=2)) %>%
    mutate(pvalue=formatC(pvalue, format = "e", digits = 2)) %>%
    mutate(padj=formatC(padj, format = "e", digits = 2)) %>%
    dplyr::select(-lfcSE,-stat) -> res.de
res.de %>% filter(log2FoldChange > 0) %>% 
    arrange(-log2FoldChange) %>%
    mutate(log2FoldChange = formatC(log2FoldChange,format="e",digits=2)) ->
    res.de.up
res.de %>% filter(log2FoldChange < 0) %>% 
    arrange(log2FoldChange) %>%
    mutate(log2FoldChange = formatC(log2FoldChange,format="e",digits=2)) ->
    res.de.down
```

```{r include=FALSE}
sd.up <- SharedData$new(res)
```

Overview
=====================================

### 

```{r include=FALSE}
comparison <- snakemake@wildcards[['comparison']]
params <- snakemake@config[['comparisons']][[comparison]]
bad.samples.data <- params[['bad-samples']]
bad.sample.string.list <- lapply(names(bad.samples), function(bs) paste0("- ",bs, ' (', bad.samples.data[bs],')'))
names(bad.sample.string.list) <- names(bad.samples)
bad.sample.string <- paste0(bad.sample.string.list,collapse='\n')
```

- Comparison: `r params[['long-name']]`
- Design: `r params[['design']]`
- Comparison attribute: `r params[['attribute']]`
- **Positive log fold changes** indicate that expression is higher in **`r paste0(params[['attribute']], '=', params[['group1_plus']])`** than in **`r paste0(params[['attribute']], '=', params[['group2_minus']])`**
- **Negative log fold changes** indicate that expression is higher in **`r paste0(params[['attribute']], '=', params[['group2_minus']])`** than in **`r paste0(params[['attribute']], '=', params[['group1_plus']])`**
- Gene prefilter function: ``r params[['prefilter']]`` (`r params[['prefilter_description']]`)
- Genes are considered significantly differentially expressed when the adjusted p value (`padj`, or equivalently, False Discovery Rate) is less than `r alpha`. 
- `DESeq2` version: `r packageVersion('DESeq2')`
- `ggp.rnaseq` version: `r packageVersion('ggp.rnaseq')`
- Report generated on `r format(Sys.time(), "%a %b %d %X %Y")`

Samples removed from analysis:

`r bad.sample.string`

### Samples included

```{r echo=FALSE}
dds <- readRDS(paste0('../',snakemake@input[['dds']]))
DT::datatable(as.data.frame(colData(dds)),width='60%',extensions=c('Scroller'),options=list(scrollY=200))
```


Higher in `r de.group.1`
===================================== 

```{r echo=FALSE}
DT::datatable(res.de.up,rownames=F,height = '100%',extensions = c('Scroller', 'Buttons'),
   options = list( scrollY = 200,  scroller = TRUE,  dom = "Bfrtip",
     buttons = c('copy', 'csv')))
```

Lower in `r de.group.1`
===================================== 

```{r echo=FALSE}
DT::datatable(res.de.down,rownames=F,height = '100%',extensions = c('Scroller', 'Buttons'),
   options = list( scrollY = 200,  scroller = TRUE,  dom = "Bfrtip",
     buttons = c('copy', 'csv')))
```

Volcano
===================================== 

```{r echo=FALSE}
my.colors <- c("Yes"="red","No"="black")
res %>%
    mutate(significant = ifelse(!is.na(padj) & padj < 0.1,"Yes","No")) %>%
    plot_ly(x=~log2FoldChange,y=~-log10(pvalue),
            color=~significant,colors=my.colors,
            type="scatter", mode="markers", hoverinfo="text", text = ~paste(SYMBOL, '<br>', GENENAME))
```

Scatterplot
===================================== 

```{r include=FALSE}
entrez.from.GOIDs <- function(go.id){
    AnnotationDbi::select(org.Mm.eg.db,key=go.id,keytype='GOALL',columns=c('ENTREZID')) %>%
    .$ENTREZID %>% unique
}
entrez.from.KEGG <- function(kegg.id){
    query <- keggGet(kegg.id)
    messy.results <- query[[1]]$GENE
    entrez <- messy.results[messy.results %in% AnnotationDbi::keys(org.Mm.eg.db,keytype='ENTREZID')]
    entrez
}
```


```{r echo=FALSE}
as.data.frame(counts(dds,normalize=TRUE)) %>%
    rownames_to_column(var="ENTREZID") %>%
    gather(-ENTREZID,key='sample',value='counts') -> df.long
data.frame(df.long,as.data.frame(colData(dds))[df.long$sample,]) -> df.long
df.long %>% group_by_(params[['attribute']],'ENTREZID') %>%
    summarize(avg = mean(counts)) %>% 
    spread(key=params[['attribute']],value='avg') -> df.mean
gene.info <- suppressMessages(AnnotationDbi::select(org.Mm.eg.db,keys=df.mean$ENTREZID,keytype='ENTREZID',columns=c('SYMBOL','GENENAME')))
suppressMessages(inner_join(df.mean,gene.info)) -> df
df$DE <- "No"
df$DE.long <- "No"
df[df$ENTREZID %in% res.de.up$ENTREZID,'DE'] <- "Up"
df[df$ENTREZID %in% res.de.down$ENTREZID,'DE'] <- "Down"
df[df$ENTREZID %in% res.de.up$ENTREZID,'DE.long'] <- paste0("Higher in ",params[['group1_plus']])
df[df$ENTREZID %in% res.de.down$ENTREZID,'DE.long'] <- paste0("Lower in ",params[['group1_plus']])
df[,'x'] <- log10(df[,params[['group2_minus']]] + 0.5)
df[,'y'] <- log10(df[,params[['group1_plus']]] + 0.5)
df[,'x.not.logged'] <- df[,params[['group2_minus']]]
df[,'y.not.logged'] <- df[,params[['group1_plus']]]
# de.colors=c("No"="black","Up"="red","Down"="blue")
# formula.x <- as.formula(paste0('~log10(',params[['group2_minus']],')'))
# formula.y <- as.formula(paste0('~log10(',params[['group1_plus']],')'))
# xvalues <- 
# df %>%
#     plot_ly(
#         x=formula.x,y=formula.y,
#         color=~DE,colors=de.colors,
#         type="scatter", mode="markers", hoverinfo="text",text = ~paste(SYMBOL, '<br>', GENENAME)
#     )
xmin <- min(df$x) - 0.5
xmax <- 1.05*max(df$x)
ymin <- min(df$y) - 0.5
ymax <- 1.05*max(df$y)

pal <- c("black","red","blue")
pal <- setNames(pal,c("No","Up","Down"))

p <- plot_ly(df,type = 'scatter', mode='markers') %>%
    add_trace(x=~x, y=~y,
              #color=~DE,colors=pal, #can't seem to get this to work
              text=~paste0(SYMBOL,'<br>',
                          GENENAME,'<br>',
                          'Mean normalized counts in ',params[['group1_plus']],": ",round(y.not.logged,digits=1),"<br>",
                          'Mean normalized counts in ',params[['group2_minus']],": ",round(x.not.logged,digits=1),"<br>",
                          "Differentially expressed: ",DE.long
                         ),
              name='All genes',
              marker=list(opacity=0.2,color='black'),
              hoverinfo='text'
             ) -> p

for(obj in snakemake@config[["gene_lists"]]){
    if(obj$type=="GO"){
        entrez <- entrez.from.GOIDs(obj$id)
        df.filtered <- df[df$ENTREZID %in% entrez,]
        p %>%
            add_trace(
                data=df.filtered,
                x=~x, y=~y,
                text=~paste0(SYMBOL,'<br>',
                             GENENAME,'<br>',
                             'Mean normalized counts in ',params[['group1_plus']],": ",round(y.not.logged,digits=1),"<br>",
                             'Mean normalized counts in ',params[['group2_minus']],": ",round(x.not.logged,digits=1),"<br>",
                             "Differentially expressed: ",DE.long
                ),
                name=obj$name,
                marker=list(color=obj$color),
                hoverinfo='text'
            ) -> p
    } else if(obj$type=="KEGG"){
        entrez <- entrez.from.KEGG(obj$id)
        df.filtered <- df[df$ENTREZID %in% entrez,]
        p %>%
            add_trace(
                data=df.filtered,
                x=~x, y=~y,
                text=~paste0(SYMBOL,'<br>',
                             GENENAME,'<br>',
                             'Mean normalized counts in ',params[['group1_plus']],": ",round(y.not.logged,digits=1),"<br>",
                             'Mean normalized counts in ',params[['group2_minus']],": ",round(x.not.logged,digits=1),"<br>",
                             "Differentially expressed: ",DE.long
                ),
                name=obj$name,
                marker=list(color=obj$color),
                hoverinfo='text'
            ) -> p
    }
}

p %>%
    layout(
        xaxis= list(
            range = c(xmin,xmax),
            title = paste0('log_10(Mean normalized counts in ', params[['group2_minus']] ,' + 0.5)')
        ),
        yaxis = list(
            range = c(ymin,ymax),
            title = paste0('log_10(Mean normalized counts in ', params[['group1_plus']] ,' + 0.5)')
        )
    ) -> p
p
```

