---
title: "fgsea"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r include=FALSE}
library(flexdashboard)
library(tidyverse)
library(DT)
library(org.Mm.eg.db)
```

```{r include=FALSE}
comparison <- snakemake@wildcards[['comparison']]

params <- snakemake@config[['comparisons']][[comparison]]
bad.samples <- params[['bad-samples']]
new.design <- as.formula(params[['design']])
prefilter.string <- params[['prefilter']]
sample.subset <- params[['subset']]
de.attribute <- params[['attribute']]
de.group.1 <- params[['group1_plus']]
de.group.2 <- params[['group2_minus']]

alpha <- 0.1

bad.samples.data <- params[['bad-samples']]
bad.sample.string.list <- lapply(names(bad.samples), function(bs) paste0("- ",bs, ' (', bad.samples.data[bs],')'))
names(bad.sample.string.list) <- names(bad.samples)
bad.sample.string <- paste0(bad.sample.string.list,collapse='\n')
```

```{r include=FALSE}
convert.entrez <- function(entrez){
    paste(suppressMessages(AnnotationDbi::mapIds(org.Mm.eg.db,keys=entrez,keytype='ENTREZID',column='SYMBOL')),collapse=',')
}
```


```{r include=FALSE}
res.fgsea <- readRDS(paste0('../',snakemake@input[['rds']]))
res.fgsea %>% filter(padj < 0.1) -> res.fgsea.sig

res.fgsea.sig %>% filter(padj < 0.1) %>% filter(NES < 0) %>% arrange(NES) %>% mutate(leadingEdge = map(leadingEdge,convert.entrez)) -> res.fgsea.down
res.fgsea.sig %>% filter(padj < 0.1) %>% filter(NES > 0) %>% arrange(-NES) %>% mutate(leadingEdge = map(leadingEdge,convert.entrez)) -> res.fgsea.up
```

Overview
===================================== 

`fgsea` is based on the following comparison:

- Comparison: `r params[['long-name']]`
- Design: `r params[['design']]`
- Comparison attribute: `r params[['attribute']]`
- **Positive log fold changes** indicate that expression is higher in **`r paste0(params[['attribute']], '=', params[['group1_plus']])`** than in **`r paste0(params[['attribute']], '=', params[['group2_minus']])`**
- **Negative log fold changes** indicate that expression is higher in **`r paste0(params[['attribute']], '=', params[['group2_minus']])`** than in **`r paste0(params[['attribute']], '=', params[['group1_plus']])`**
- Gene prefilter function: ``r params[['prefilter']]`` (`r params[['prefilter_description']]`)
- `DESeq2` version: `r packageVersion('DESeq2')`
- `ggp.rnaseq` version: `r packageVersion('ggp.rnaseq')`
- Report generated on `r format(Sys.time(), "%a %b %d %X %Y")`

Samples removed from analysis:

`r bad.sample.string`

`fgsea` parameters:

- Minimum pathway size: `r snakemake@config[['fgsea']][['minSize']]` genes
- Maximum pathway size: `r snakemake@config[['fgsea']][['maxSize']]` genes
- Number of permutations: `r as.character(snakemake@config[['fgsea']][['nperm']])`
- `fgsea` version: `r packageVersion('fgsea')`

Higher in `r de.group.1`
===================================== 

```{r echo=FALSE}
DT::datatable(res.fgsea.up,rownames=F,height = '100%',extensions = c('Scroller', 'Buttons'),
   options = list( scrollY = 200,  scroller = TRUE,  dom = "Bfrtip",
     buttons = c('copy', 'csv')))
```

Lower in `r de.group.1`
===================================== 

```{r echo=FALSE}
DT::datatable(res.fgsea.down,rownames=F,height = '100%',extensions = c('Scroller', 'Buttons'),
   options = list( scrollY = 200,  scroller = TRUE,  dom = "Bfrtip",
     buttons = c('copy', 'csv')))
```
