---
title: "Heatmap"
output: 
  flexdashboard::flex_dashboard
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(ggp.rnaseq)
library(DESeq2)
library(DT)
library(crosstalk)
library(d3scatter)
library(plotly)
library(org.Mm.eg.db)
library(heatmaply)
```

```{r include=FALSE}
heatmap.name <- snakemake@wildcards[['heatmap']]
params <- snakemake@config[['heatmaps']][[heatmap.name]]
sample.attributes <- params[['sample_attributes']]
n.top.genes <- params[['number_top_genes']]
```

```{r include=FALSE}
vsd <- readRDS(paste0('../',snakemake@input[['vsd']]))
```

`r heatmap.name`
=====================================

```{r include=FALSE}
ord <- order(-rowVars(assay(vsd)))
m.top.200 <- assay(vsd)[ord[1:200],]
rownames(m.top.200) <- AnnotationDbi::mapIds(org.Mm.eg.db,keys=rownames(m.top.200),keytype='ENTREZID',column='SYMBOL')
```

### Heatmap

```{r}
sample.data <- as.data.frame(colData(vsd))[,sample.attributes,drop=FALSE]
heatmaply(m.top.200,scale='row',col_side_colors = sample.data,margins=c(100,100))
```

### Info

- Heatmap is based on the `r n.top.genes` genes with the highest variance after applying the `DESeq2` function `varianceStabilizingTransformation`.
- Gene prefilter function: ``r params[['prefilter']]`` (`r params[['prefilter_description']]`)
- `DESeq2` version: `r packageVersion('DESeq2')`
- `ggp.rnaseq` version: `r packageVersion('ggp.rnaseq')`
- Report generated on `r format(Sys.time(), "%a %b %d %X %Y")`